### 2022-11-09 manuscript/manuscript.rmd
Adding new analyses

---
title: Analyzing COVID-19 Mortality at the County Level in the U.S. Using Generalized Additive
  Models
author: "Christian Testa"
date: '`r Sys.Date()`'
output: pdf_document
abstract: this is an abstract
---

```{r setup, echo = FALSE}
library(here)

knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  out.width = '100%'
)
```

# Introduction

As we enter the third winter with novel coronavirus disease (COVID-19) in the
United States (U.S.), evidence reflecting intense disparities in COVID-19 mortality rates 
among socially advantaged and disadvantaged populations persists.  Eliminating
inequities in health outcomes is a major policy goal of the Biden
administration, representing an revitalized focus on health justice
and underlining the importance of adequate data reporting systems that measure
the prevalent health inequities.  To this end, generalized additive models
(GAMs) provide a flexible regression framework which can illustrate the evolving
roles and relationships social conditions have with trends in COVID-19
mortality.

# Background

Having passed over 1 million COVID-19 deaths in the United States earlier this
year, and facing uncertain prospects for the third COVID-19 winter looming even
as new iterations on the COVID-19 vaccines become available, it remains critical
that inequities in COVID-19 outcomes are documented and analyzed to reckon with
the unjust and unfair burden of preventable illness.  Even though the first
vaccines were granted emergency use authorization by the U.S. Food and Drug
Administration in 2020, with the first shots going in arms in December 2020,
COVID-19 is still continuing to cause hundreds of deaths a day in the U.S. in
the late summer and early fall of 2022.  The new bivalent vaccines released at
the end of August 2022 contain messenger RNA components of the original strain
as well as the BA.4 and BA.5 lineages in an effort to make the nation's immunity
more up-to-date and robust against the myriad of phylogenetic directions the
COVID-19 virus is exploring.  Despite the updated bivalent boosters representing
a significant step forward in prevention strategy, less than 4% of eligible
Americans had taken the booster in the first month after it became available. As
such, and with an enduring history of inequities in health care availability, it
is not clear that all communities will be equally able to take advantage of the
new vaccines and inequities in COVID-19 illness and mortality may persist
despite the technological innovations in vaccine therapy.

```{r figure 1, fig.cap = "Monthly estimates of COVID-19 mortality per 100,000 residents by state are shown.  States are in distinct colors to make each state's trajectory visible. Reference the online version TODO: XXX"}
knitr::include_graphics(here("analysis/01_state_trajectories/state_covid19_mortality_trajectories.png"))
```

# Methods

## Data Sources

The following variables were retrieved at the county level:

  * Counts of deaths by month with underlying cause of death as
  COVID-19 (ICD-10 code U07.1) from CDC WONDER, Provisional Multiple Cause of Death
  Data
  * Population Size Estimates for 2020 from the U.S. Census (Redistricting 
  File Public-Law 94-171 Dataset). *Note:*  While the complete data for the 2020
  U.S. Census have not yet been released, the summary tables including
  population counts have been released as mandated by Congress for the purposes
  of redistricting and updating the boundaries of the 435 voting Congressional
  Districts, which are what is made us of here as population size estimates.
  * Percent Black, percent Hispanic, percent Asian, percent American Indian or
  Alaska Native, Percent below the federal poverty line, median household
  income, median age, percent 65 years or older, and percent 75 years or older
  from the 2014-2019 5-year American Community Survey. 
  *Note:* while more recent American Community Survey 5-year estimates are
  available, survey data collection in 2020 was severely affected by the
  pandemic with the 2020 data release denoted as "experimental".

## Generalized Additive Models

# Results

We used GAMs to describe the following:

  * the changing associations between individual area-based measures
  and COVID-19 mortality over time;
  * the changing associations between bivariate area-based measures
  and COVID-19 mortality over time; 
  * associations with area-based measures and COVID-19 mortality after
  controlling for spatio-temporal autocorrelation.

Additionally, as sensitivity analyses, random forest models were fit to compare
with the coefficient estimates from the GAMs.

# Discussion

## Limitations

  * Missing data 
  * Ecological fallacy 
  * Age effects
  * Fixed population denominators -> real-time denominator estimates
  * Vaccination and mobility effects

# Conclusions

# References

<!-- 1 million covid deaths: 
https://coronavirus.jhu.edu/from-our-experts/u-s-officially-surpasses-1-million-covid-19-deaths -->

<!-- nearly 225,000 Americans have died since the start of 2022
https://abcnews.go.com/Health/225000-americans-died-covid-19-start-2022-viral/story?id=90339579 -->

<!-- 
vaccine timeline
https://www.mayoclinic.org/coronavirus-covid-19/history-disease-outbreaks-vaccine-timeline/covid-19
-->

<!--
less than 4% of americans had gotten the booster 1 month into availability
https://www.nbcnews.com/health/health-news/updated-covid-booster-shots-doses-administered-cdc-rcna48960
-->

<!-- 2020 ACS data are experimental 
https://www.census.gov/newsroom/press-releases/2021/experimental-2020-acs-1-year-data.html -->

<!--
Spatial-temporal generalized additive model for modeling COVID-19 mortality risk in Toronto, Canada
https://www.ncbi.nlm.nih.gov/pmc/articles/PMC8257405/ -->

<!-- 
Using GAM functions and Markov-Switching models in an evaluation framework to assess countries’ performance in controlling the COVID-19 pandemic
https://bmcpublichealth.biomedcentral.com/articles/10.1186/s12889-021-11891-6 -->

<!--
Statistical modelling of COVID-19 data: Putting generalized additive models to work
https://journals.sagepub.com/doi/full/10.1177/1471082X221124628
-->

<!--
Inference following multiple imputation for generalized additive models: an investigation of the median p-value rule with applications to the Pulmonary Hypertension Association Registry and Colorado COVID-19 hospitalization data
https://bmcmedresmethodol.biomedcentral.com/articles/10.1186/s12874-022-01613-w
-->

<!--
Generalized Additive Models to Capture the Death Rates in Canada COVID-19
https://link.springer.com/chapter/10.1007/978-3-030-85053-1_7 -->

<!--
Contribution of Two-Dose Vaccination Toward the Reduction of COVID-19 Cases, ICU Hospitalizations and Deaths in Chile Assessed Through Explanatory Generalized Additive Models for Location, Scale, and Shape
https://www.frontiersin.org/articles/10.3389/fpubh.2022.815036/full -->

<!-- 
Combining graph neural networks and spatio-temporal disease models to improve the prediction of weekly COVID-19 cases in Germany
https://www.nature.com/articles/s41598-022-07757-5 -->

<!-- 
Association between mobility patterns and COVID-19 transmission in the USA: a mathematical modelling study
https://www.sciencedirect.com/science/article/pii/S1473309920305533 -->

<!-- 
Examining spatiotemporal evolution of racial/ethnic disparities in human mobility and COVID-19 health outcomes: Evidence from the contiguous United States
https://www.sciencedirect.com/science/article/pii/S2210670721007721 -->

<!--
Social, economic, and environmental factors influencing the basic reproduction number of COVID-19 across countries
https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0252373 -->

<!-- 
https://asbates.rbind.io/2019/05/03/gams-for-time-series/
-->

<!--
https://environmentalcomputing.net/statistics/gams/
-->

<!--
https://petolau.github.io/Analyzing-double-seasonal-time-series-with-GAM-in-R/
-->

# Supplementary Materials