### 2022-11-23 manuscript/manuscript.rmd
Working on the manuscript

---
title: Documenting Social, Geographic, and Economic Inequities in COVID-19 Mortality at the County Level in the US Using Generalized Additive Models
author: 
  - name: "Christian Testa"
    institute: "1"
    email: "ctesta@hsph.harvard.edu"
    correspondence: true
institute: 
  - "1": "Department of Social and Behavioral Sciences, Harvard T.H. Chan School of Public Health"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  bookdown::pdf_document2:
    pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
    toc: false
    number_sections: false
abstract: |
  We present three types of applications of generalized additive models (GAMs) to
  COVID-19 mortality rates in the US for the purpose of advancing methods to
  document inequities with respect to which communities suffered disproportionate
  COVID-19 mortality rates. First, GAMs can be used to describe the changing
  relationship between COVID-19 mortality and county-level covariates
  (sociodemographic, economic, and political metrics) over time. Second, GAMs can
  be used to perform spatiotemporal smoothing that pools information over time and
  space so that county level mortality rates that tend to be noisy due to small
  population counts or stochasticity can be summarized by a smooth, dynamic latent
  surface describing the mortality risk associated with geographic locations over
  time. Third, estimation of the associations between county-level covariates and
  COVID-19 mortality after controlling for spatiotemporal effects allow for the
  distinguishing of what patterns in COVID-19 mortality were more plausibly due to
  geography than variation in county-level covariates. Each of these approaches
  provides a valuable approach and perspective to documenting inequities in
  COVID-19 mortality by addressing the question of which populations have suffered
  the worst burden of COVID-19 mortality taking into account the nonlinear
  spatial, temporal, and social patterning of disease.
  \newline

  \noindent Abbreviations used: United States (US), Coronavirus Disease 2019 (COVID-19), Generalized
  Additive Model (GAM), Centers for Disease Control and Prevention (CDC)
bibliography: bibliography.bib
csl: apa.csl
link-citations: true
urlcolor: blue
linkcolor: black
header-includes: 
  - \usepackage{pdflscape}
  - \newcommand{\blandscape}{\begin{landscape}}
  - \newcommand{\elandscape}{\end{landscape}}
---

<!--
TODO:

Throughout/Overarching: 
  - Do I need to update all the models to separately control for age? Probably.
  - Do I want to have a supplementary materials section?
  - Talk about ecosocial theory (?) (Who benefits from harm done to whom?)

Abstract:
  - Add link to GitHub repository

Intro: 
  - add geographic background, like where the first cases were reported, 
  how COVID exploded in NYC, then the surges in the Midwest and South
  - add a primer on the relevance of spatiotemporal modeling

Methods: 
  - add intro to ICEraceinc (methods)
  - describe negative binomial models

Results:
  - add figure for temporal two-way interaction effects
  - present rate ratios (results)
  - present model comparisons (results, F-tests, AIC)

Discussion:
  - talk about how COVID has not been a win for health equity (discussion)
  - talk about limitations to being causal
  - talk about future directions

Conclusion:
  - summarize
-->

```{r setup, echo = FALSE}
suppressMessages(library(here))

knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  out.width = '100%'
)
```

# Introduction

As we enter the third winter with the novel coronavirus disease 
COVID-19 in the United States, evidence documenting the intense disparities in
COVID-19 mortality rates comparing socially advantaged and disadvantaged
populations continues to mount. Eliminating inequities in health outcomes has
been stated as a major policy goal of the Biden administration
[@the_white_house_executive_2021; @the_white_house_advancing_2022], representing
a revitalized focus on health justice and underlining the importance of adequate
data reporting systems that measure the prevalent health inequities. To this
end, we use generalized additive models (GAMs) as a flexible regression
framework to illustrate the evolving roles and relationships sociodemographic,
geographic, and economic conditions have with respect to trends in COVID-19
mortality. The code, data, and documentation necessary to reproduce the analyses
contained in this paper are online and free to access at <https://github.com/ctesta01/covid.gradient.estimation>.

### Background

Having passed over 1 million COVID-19 deaths in the United States earlier this
year [@donovan_us_2022], and facing uncertain prospects for the third COVID-19
winter looming even as new iterations on the COVID-19 vaccines become available,
it remains critical that inequities in COVID-19 outcomes are documented and
analyzed to reckon with the unjust and unfair burden of preventable illness.
Even though the first vaccines were granted emergency use authorization by the
U.S. Food and Drug Administration in 2020 [@mayo_clinic_history_2022], with the
first shots going in arms in December 2020, COVID-19 is still continuing to
cause hundreds of deaths a day in the US in the fall of 2022
[@noauthor_united_2022].  The new bivalent vaccines released at the end of
August 2022 contain mRNA sequences from both the original strain as well as
the recently emergent BA.4 and BA.5 lineages in an effort to make the nation's immunity more
up-to-date and robust against the myriad of phylogenetic directions the COVID-19
virus is evolving to explore [@office_of_the_commissioner_coronavirus_2022].
Despite the updated bivalent boosters representing a significant step forward in
prevention strategy, less than 4% of eligible Americans had taken the booster in
the first month after it became available [@bendix_less_2022; @lambert_most_2022]. 
As such, and with an enduring history of inequities in health care access 
in the US [@ortega_ending_2021; @blendon_inequities_2002; @bailey_how_2021;
@okonkwo_covid-19_2021; @rapp_statelevel_2022;
@chrisler_ageism_2016; @feldman_health_2021; @carpenter_health_2021;
@whitehead_outness_2016], it is clear that without further intervention not all
communities will be equally able to benefit from the updated vaccines and
inequities in COVID-19 illness and mortality may persist despite the
technological innovations in vaccine technology.

### The Role of Geography in COVID-19

```{r figure-county-trends, fig.cap = "Estimates of monthly COVID-19 mortality rates per 100,000 person-years by county organized by Census Division.  For each division, the median trendline and quantile ranges are shown weighted by county population size.", out.width = "6in", fig.align = 'center', fig.env = "figure*"}
knitr::include_graphics(here("analysis/02_plot_county_trends/county_trends_by_division.png"))
```

Prior literature has demonstrated that geography has played a significant role
in the spread and impact of of COVID-19. Methods employed to highlight the role
geography plays have included quantile regression [@sigler_socio-spatial_2021],
Besag-York-Molli\'e mixed models [@whittle_ecological_2020], spatial cluster analysis
[@sugg_mapping_2021], geographically weighted regression [@mollalo_gis-based_2020; @park_covid-19_2021], and others.

Figure \@ref(fig:figure-county-trends) shows the monthly COVID-19 mortality
rates for the counties grouped within each of the nine U.S. Census Divisions
[@us_department_of_commerce_economics_and_statistics_administration_census_2000]. 
The figure summarizes each division's median mortality rates
weighted by county population size. Quantile ranges are included to illustrate
the range of mortality rates observed, again weighted by county population size.
Weighting based on county population size is used as opposed to equal weighting
for calculating the median and quantile ranges so that in divisions where low
population counties predominate, mortality rate estimates that are noisy due to
small population counts are not weighted equally to more reliable mortality rate
estimates from larger population counties. Notably, the mortality associated
with the early surge of cases starting in NYC and spreading through New York,
New Jersey, and Massachusetts is visible in the Middle Atlantic and New England
division figures. The figure also illustrates how the first peak in the
mortality time-series for states in the Midwest (West North Central, East North
Central) occurred later, in late 2020 and going into early 2021.

In the US context, one of the key aspects to the geographic story of COVID-19's
spread and diffusion was the early surge of cases and epicenter in New York City
during March 2020 [@thompson_covid-19_2020] followed by subsequent waves of cases in the
South and Midwest [@shumaker_covid-19_2020; @scott_these_2020; @glenza_covid_2020]. As Park et al. 
stated summarized trends in the US from March 2020 to May 2021, "hot spots have shifted from densely populated cities and the states with a high percentage of socially vulnerable individuals to the states with relatively relaxed social distancing requirements, and then to the states with low vaccination rates" [-@park_covid-19_2021].

When considering the drivers of the COVID-19 pandemic, it's necessary to note that 
geography and social conditions are inextricably linked. In July 2021, the CDC
reported that "the COVID-19 cumulative death rate in non-metropolitan areas has
exceeded that of metropolitan areas since December 2020," noting that of the
approximately 1/5th of Americans who live in rural areas, many "are considered
highly vulnerable according to CDC’s Social Vulnerability Index (SVI), which
includes factors such as housing, transportation, socioeconomic status, race,
and ethnicity" [@cdc_location_2021]. Moreover, rural communities often have
lower health insurance rates, higher disability rates, older populations, and
limited access to health care. One of CDC's Morbidity and Mortality Weekly
Reports found that vaccination against COVID-19 was lower in rural communities
than in urban communities between December 2020 and April 2021
[@murthy_disparities_2021].

### The Social Determinants of COVID-19 Mortality

Even since the beginning of the COVID-19 outbreak in the US, data reflected
sharp inequities in mortality rates.  During January 22nd to May 5th 2020,
county COVID-19 mortality rates were 4.94 (95% CI 4.78, 5.09) times higher in
counties in the highest quintile of percent People of Color (61%-100%) compared
to counties in the lowest quintile of percent People of Color (0%-17.2%)
[@chen_revealing_2021].  This was not wholly unanticipated: as COVID-19 was
beginning to take off in the US, some were already calling attention to the fact
that COVID-19 threatened to exacerbate existing disparities
[@kim_covid-19_2020]. Kim, Marrast, and Conigliaro noted three structural
barriers in COVID-19 prevention and care: 1) originally requiring New York
residents to have a doctor's prescription for a COVID-19 test reduced the
opportunity for healthcare for People of Color as they are less likely to have a
primary care provider; 2) drive-through testing made testing disadvantaged those
who rely on public transportation; 3) quarantining at home while waiting the
7-10 days originally required for COVID-19 test results to come back posed an
economic and social challenge that many in already financially difficult
situations may not have been able to take on [-@kim_covid-19_2020]. Others noted
yet more reasons why COVID-19 threatened to worsen an already inequitable
healthcare landscape in the US: those who reside in prisons and jails,
immigrants and undocumented people, people with disabilities, and people
experiencing homeless all face additional challenges in seeking and getting the
healthcare they deserve [@okonkwo_covid-19_2021]. It's not just a hypothetical 
that those with more structurally enfranchised privileges have been more able to
mitigate their risk of negative health outcomes associated with COVID-19 
— during February 1 to April 1, 2020, New York City residents who lived in more affluent 
neighborhoods were more likely to have left the city, while New York City residents
from more Black and Hispanic neighborhoods were more likely to continue working [@coven_disparities_2020].

> Contrary to the oft used phrase that the ‘virus does not discriminate’, the
data presented here suggest that this virus, as many other infectious diseases,
has the greatest implications for the most vulnerable people. The intersections
between health and human rights are clear—the health of a society and
vulnerability to a pandemic are directly related to its human rights track
record for those who are marginalised. [@okonkwo_covid-19_2021]

When vaccines became available, vaccine appointments were often only available
to be scheduled through online web-portals contributing to the inequities
between those who had internet access and technological literacy and those who
didn't [@press_inequities_2021]. Vaccination sites have not been equally
distributed and areas determined to be vaccine deserts have been found to have
disproportionately more Black and Hispanic residents [@rader_spatial_2021]. In
fact, healthcare facilities in counties with higher Black composition had 32%
(95% CI 14%-47%) lower odds of serving as vaccine sites
[@hernandez_disparities_2022].

What vaccination has been administered hasn't suddenly erased the unequal burden
of COVID-19 either; in August 2022 the New York Times was reporting "Black death
rates at this winter’s peak were greater than those of white people by 34
percent in rural areas, 40 percent in small or medium cities and 57 percent in
big cities and their suburbs" [@goldstein_during_2022]. As COVID-19 case and
mortality rates have waxed and waned, the inequities have widen and shrunken,
often with racial/ethnic inequities growing during times when COVID-19 rates
have surged [@hill_covid-19_2022]. During 2022, the age-standardized
COVID-19 mortality rates for white people have, at times, been slightly 
higher than those of Black and Hispanic people, predominantly because the 
mortality rate among white people has increased [@johnson_whites_2022]. 
It's important to note that white COVID-19 mortality rates overtaking 
the Black mortality rates does not imply that equity has been established: neither
does this undo the cumulative impact of mortality (which has been twice as high for
Black people compared to white people [@hill_covid-19_2022]), nor does it imply 
the underlying systemic barriers to equity have been overturned
[@del_rios_covid-19_2022]. As Del Rios, Chomilo, and Lewis note, instead,
COVID-19 leaves in its wake more years of life expectancy lost, wages lost, and
degradation of mental health in Communities of Color.

# Methods

## Data Sources

The following variables were retrieved at the county level:

  * Counts of COVID-19 deaths [@the_new_york_times_coronavirus_2021].
  * Population size estimates for 2020 from the U.S. Census (Redistricting 
  File Public-Law 94-171 Dataset) [@us_census_bureau_2020_2021].
  * Median age, median household income, racial/ethnic composition, population
  density, percent below the federal poverty line, and number of households with
  high (\$100k+)/low (<\$25k) household income by racial/ethnic group from the
  2014-2019 5-year American Community Survey [@us_census_bureau_american_2020]
  through the `tidyverse` R package [@walker_tidycensus_2022]. 
  * Votes cast in the 2020 presidential election [@mit_election_data_and_science_lab_county_2022]

## Generalized Additive Models

Traditional generalized linear models fit a regression model using the
functional relation $g(\mu_i) = \mathbf X_i \beta$ where Examples of common link
functions include the identity function for linear regression, $\log$ for
Poisson regression, and the $\text{logit}$ function for logistic regression.

In contrast, generalized additive models (GAMs) improve upon generalized linear models by
allowing for the fitting of smooth functions that transform the $x$ variables.
This is a convenient means to include nonlinear relationships between the
outcome and predictor variables. Whereas a generalized linear model may have looked like 
$$ g(\mu_i) = \beta_0 + \beta_1 x_{i1} + \beta_2 x_{i2} + \beta_3 x_{i3} ...$$ 
a generalized additive model could look like 
$$ g(\mu_i) = \mathbf A_i \mathbf \theta + f_1(x_{i1}) + f_2(x_{i2}) + f_3(x_{i3}, x_{i4}) + \dots$$
where $\mu_i \equiv \mathbb E(Y_i)$, $\mathbf A_i$ is a row of the model matrix 
for any strictly non-parametric model components, $\mathbf \theta$ is the corresponding
parameter vector, and the $f_j$ are smooth functions of the covariates $x_k$ [@wood_generalized_2017].

```{r gam-demo-figure, fig.cap = "A demonstration of fitting regression models using B-spline basis functions. Part A. shows the simulated dataset. Part B. shows the B-spline basis functions. Part C. shows the weighting of the B-spline basis functions and their weighted sum with an intercept added to create the regression model fit to the data."}

knitr::include_graphics(here("analysis/10_supplementary_gam_figures/gam_demo_figure.png"))
```

The smooth functions estimated as part of fitting a GAM are constructed using
spline basis functions. These spline basis functions allow for the smooth
interpolation of trends in the data allowing for the incorporation of
nonlinearity. To avoid overfitting the data, a penalty term is introduced that
controls the degree of "wiggliness" or smoothness and this penalty term is fit
using generalized cross validation. B-splines are one kind of spline basis function 
that is commonly used and are especially popular because of their property that they 
are non-negative on only a finite interval.

While a variety of varying kinds of spline bases exist (cubic splines, B- and
P-splines, thin-plate splines, etc.), our particular application setting
warrants the use of tensor-product smooths because we require anisotropic
penalization. That is to say, when we seek to create spatiotemporally smoothed
model estimates, it's inappropriate to assume that the amount of smoothing
across space should be the same as the amount of smoothing across time because
they're in fundamentally different units (space being measured in Cartesian
coordinates and time being measured in years, months, days, etc.). More details
about tensor-product smooths are available in Simon Wood's *Generalized Additive Models - An Introduction with R* [-@wood_generalized_2017].

## Overdispersion and Event Modeling

Above and beyond using the GAM framework to allow for flexible, nonlinear
relationships between our observed county-level variables and COVID-19
mortality, we must have a model specification that agrees with the data
generating process. In our case, the data generated are counts of deaths per
population, and the class of models most suited to represent counts of events
are Poisson, quasi-Poisson, and negative binomial models. Here we've chosen to 
use the negative binomial model as it accounts for overdispersion and has 
the intuitive interpretation of a Poisson model with gamma distributed underlying
rate parameter [@gelman_bayesian_2013].

A negative binomial model in the context of modeling count data can be written
$$y_i \sim \text{NegativeBinomial}(u_i \theta_i, \phi),$$
where $y_i$ are the counts observed, $u_i$ is the exposure, $\theta_i$ are
the rates, and $\phi$ determines the amount of overdispersion. The rates
are modeled as $\theta_i = e^{X_i \beta}$ where $X_i$ are the observed
covariates and $\beta$ are the coefficients on the covariates corresponding
to log rate ratios. The logarithm of the exposure, $\log(u_i)$ is often called 
(and later herein referenced as) the offset. In epidemiological contexts, the offset 
is often representative of the amount of person-time during which observations were recorded. Whereas the Poisson model holds that $\text{var}(y) = \mu$ where $\mu$
is the average rate, the negative binomial model instead assumes that
$\text{var}(y) = \mu + \mu^2/\phi$. The Poisson model
is a special case of the negative binomial model when $\phi \to \infty$.

An alternative and equivalent formulation of the negative binomial that is commonly 
used makes the connection to the Poisson model even more clear: instead of 
$\theta_i$ and $\phi$, using $\alpha$ and $\beta$, 
$$y \sim \text{NegativeBinomial}(\alpha, \beta), \text{ and}$$
$$\text{NegativeBinomial}(y|\alpha, \beta) = \int \text{Poisson}(y|\theta) \text{Gamma}(\theta | \alpha, \beta) d\theta.$$

Introductions to and additional exposition on the negative binomial model,
especially in the context of modeling counts of events outcome data, are
available in *Bayesian Data Analysis* and *Regression and Other Stories*
[@gelman_regression_2020; @gelman_bayesian_2013].

## Variables of Interest

Besides age, which is the greatest risk factor for COVID-19 mortality, we have
included the following variables as covariates of interest:

  * Population Density per Square Mile
  * Median Household Income
  * Proportion in Poverty
  * the Index of Concentration at the Extremes for Racialized Economic Segregation 
  * Political Lean in the 2020 Election (1 = 100% Republican votes, -1 = 100% Democratic Votes)

#### Index of Concentration at the Extremes for Racialized Economic Segregation

The Index of Concentration at the Extremes (ICE) is a measure which describes
how concentrated a given area's population is in terms of the extreme ends of
privilege and marginalization [@krieger_public_2016]. 
In general, the ICE measure is formulated as

$$ \text{ICE} = \frac{\text{Number of People in Most Privileged Category} - \text{Number of People in Least Privileged Category}}{\text{Total Population}}$$

Applying the ICE approach to a specific context involves defining the axes of
privilege of interest.  In this case, data on racialized economic segregation
are used from the US Census American Community Survey on White households
earning more than $100,000 a year (the most privileged group) or households of
People of Color earning less than $25,000 a year (the least privileged group).
This variable is referred to throughout as the ICE for racialized economic
segregation, or `ICEraceinc` in the code.

Compared with the Gini
coefficient which is one of the most popular methods for summarizing area-level
rates of inequities, the ICE the major advantage that it is more suitable for
describing inequities at smaller area levels [@krieger_public_2016]. While the
Gini coefficient measures within-area dissimilarity (as in, for example, how
unequal wealth is distributed within a county), the ICE measure establishes
where on a spectrum a given county's population falls allowing for comparison
across counties. The Gini coefficient suffers from the fact that areas which are
made up of relatively homogeneous populations will appear as having low
within-area inequality (and therefore low Gini coefficient). Instead, the ICE
measures how much of the population is either privileged or not. The Gini
coefficient remains useful for reporting on the degree of inequity in larger
areal units (like countries, states, and regions), but at smaller areal units
(like counties, ZIP codes, census tracts) can be more difficult to interpret and
compare. Maps of ICE measure can elucidate what spatial social segregation and
polarization exist, and the ICE for racialized economic segregation has been
repeatedly and significantly associated with COVID-19 outcomes
[@hanage_covid-19_2020; @chen_revealing_2021; @krieger_relationship_2022;
@eichenbaum_health_2022; @brown_ecological_2021; @sonderlund_racialized_2022;
@saha_neighborhood-level_2020].

#### Political Lean

Political lean has been associated with COVID-19 mortality in numerous studies,
with plausible mechanisms explaining the association including differences in
non-pharmaceutical intervention uptake (mask usage, social distancing,
quarantining), differences in rates of vaccination, differences in political
leadership's messaging, resource allocation, and the adoption of policy
interventions [@krieger_relationship_2022; @kaashoek_evolving_2021; @leonhardt_red_2021; @gonzalez_conservatism_2021; @grossman_political_2020].

# Results

## Application 1: Non-Spatial Covariate Effects Over Time

The GAM models are fit with the following structure using 
the `gam` function from the `mgcv` package [@wood_mgcv_2022]:

```{r gam-demo-code, echo = TRUE, eval=FALSE}
gam(
  formula = deaths ~ s(median_age) + te(covariate, date, d=c(1,1)), # regression formula
  offset = log(popsize/1e5/12), # our offset represents the person-time
  data = df,     # our data-set of county-level observations
  family = nb(link='log')  # indicates negative binomial family and a log-link function
)
```

The formula used puts a one-dimensional smoothing spline to represent a
nonlinear age-effect and a two-dimensional tensor-product smooth on the
interaction between the given covariate and the date. The `d=c(1,1)` argument
provides the instruction necessary to consider covariate and date as being on
separate scales and therefore to fit the tensor-product smooth with anisotropy —
that is, to allow for independent amounts of scaling in the dimensions of the
covariate and time. The offset used structures the regression to estimate rates
in units of person-time per 100,000 person-years. Since the death counts are
aggregated to the monthly level, the person-time in units of 100,000
person-years are calculated by taking each county's population size, dividing by
100,000, and dividing by 12 (for the 12 months in a year).

The above structure is used to estimate models for our different `covariate`
variables of interest: median income, percent in poverty, the ICE for racialized
economic segregation, political lean.  Whereas the other models treat median age
as a variable to adjust for, the model presenting the effect of median age over
time treats median age as the main covariate of interest including it as the 
`te(covariate, date, d=c(1,1))` and dropping the `s(median_age)` term which is 
redundant. Results of these models are summarized in \@ref(fig:one-variable-over-time)
where the additional COVID-19 mortality associated with each covariate is visualized.

(ref:medianage) [@us_census_bureau_nation_2022]

```{r one-variable-over-time, fig.cap = "Additional COVID-19 mortality associated with covariates over time. A) age and time, B) log (base 10) population density and time, C) median income and time, D) proportion in poverty and time, E) ICEraceinc and time, F) political lean and time. In panels B-F the effect of age is marginalized out using the median age in the US, 38.8 (ref:medianage).", fig.align = 'center', out.height = '75%'}
knitr::include_graphics(here("analysis/04_one_variable_over_time/panel_figure.png"))
```

Likelihood ratio tests confirmed that models with covariates interacted with
time had significantly lower residual deviance ($p \leq 2.2e^{-16}$ for all
models) compared to models only including a spline term on median age and the
given covariate not interacted with time. The model interacting median age 
and time was compared to a model with a spline for median age alone. Akaike Information
Criteria values were also lower for all models compared to models that did
not interact the covariates with time.

Another non-spatial application of GAMs to describe the distribution of 
COVID-19 mortality in the US over time is to consider the associations of 
mortality with three-way interactions of time and two covariates taken together. 
In the following example, the formula used is `deaths ~ te(date, ICEraceinc, median_age, d=c(1,1,1))`.  Again, the `d` argument specifying the marginal 
basis dimensions is used indicate that each of the `date`, `ICEraceinc`, and 
`median_age` measures are in different units and should not be smoothed assuming
that a one unit difference in one variable is comparable to a one unit difference
in another variable.

```{r two-variables-over-time, fig.cap="The changing interacted effect of the ICE for racialized economic segregation and age over time. A) May 2020, B) November 2020, C) May 2021, D) October 2021, E) February 2022, F) August 2022. An animated version is available online at "}
knitr::include_graphics(here("analysis/05_two_variables_at_a_time/animation_ICEraceinc_age/panel_figure.png"))
```

## Evaluating Spatial and Temporal Autocorrelation

```{r figure county neighboring graph, fig.cap = "A) The spatial neighbor graph for counties in the US. In this graph, county centroids are connected if the two counties are adjacent to one another (i.e., share a boundary). This and similar graphs that represent neighbor relationships or geographic proximity are often used in modeling contexts to create spatial weights that account for which regions help to explain a given region's observed data.", fig.show='hold', out.width = '50%', fig.env = "figure*"} 
desired_caption <- 
  " B) Moran's scatter plot \\citep{anselin_simple_1996} depicts for each region (in our case, counties) on the x-axis the region rate and on the y-axis the average of the region's neighbors' rates. This diagnostic plot provides a simple means to assess to what degree neighboring regions share similar rates. In a setting where regions have high correlation, Moran's scatter plot will exhibit a high degree of positive correlation."
knitr::include_graphics(here("analysis/08_spatiotemporal_autocorrelation/county_neighboring_relations.png")) 
knitr::include_graphics(here("analysis/08_spatiotemporal_autocorrelation/morans_plot.png"))
```

```{r figure 4, fig.cap = ""}
knitr::include_graphics(here("analysis/08_spatiotemporal_autocorrelation/temporal_acf.png"))
```

## Regression Modeling Results

We used GAMs to describe the following:

  * the changing associations between individual area-based measures
  and COVID-19 mortality over time;
  * the changing associations between bivariate area-based measures
  and COVID-19 mortality over time; 
  * associations with area-based measures and COVID-19 mortality after
  controlling for spatio-temporal autocorrelation.

Additionally, as sensitivity analyses, random forest models were fit to compare
with the coefficient estimates from the GAMs.

```{r figure 5, fig.cap = ""}
knitr::include_graphics(here("analysis/09_spatiotemporal_models/temporal_panel_figure.png"))
```

```{r figure 6, fig.cap = ""}
knitr::include_graphics(here("analysis/09_spatiotemporal_models/panel_figure.png"))
```

# Discussion

## Limitations

  * Ecological fallacy 
  * Age effects
  * Fixed population denominators -> real-time denominator estimates
  * Vaccination and mobility effects
  * Equal Smoothing Across All Areas
  * Soap Film Smoothers

# Conclusions

# References

<!-- 1 million covid deaths: 
https://coronavirus.jhu.edu/from-our-experts/u-s-officially-surpasses-1-million-covid-19-deaths -->

<!-- nearly 225,000 Americans have died since the start of 2022
https://abcnews.go.com/Health/225000-americans-died-covid-19-start-2022-viral/story?id=90339579 -->

<!-- 
vaccine timeline
https://www.mayoclinic.org/coronavirus-covid-19/history-disease-outbreaks-vaccine-timeline/covid-19
-->

<!--
less than 4% of americans had gotten the booster 1 month into availability
https://www.nbcnews.com/health/health-news/updated-covid-booster-shots-doses-administered-cdc-rcna48960
-->

<!-- 2020 ACS data are experimental 
https://www.census.gov/newsroom/press-releases/2021/experimental-2020-acs-1-year-data.html -->

<!--
Spatial-temporal generalized additive model for modeling COVID-19 mortality risk in Toronto, Canada
https://www.ncbi.nlm.nih.gov/pmc/articles/PMC8257405/ -->

<!-- 
Using GAM functions and Markov-Switching models in an evaluation framework to assess countries’ performance in controlling the COVID-19 pandemic
https://bmcpublichealth.biomedcentral.com/articles/10.1186/s12889-021-11891-6 -->

<!--
Statistical modelling of COVID-19 data: Putting generalized additive models to work
https://journals.sagepub.com/doi/full/10.1177/1471082X221124628
-->

<!--
Inference following multiple imputation for generalized additive models: an investigation of the median p-value rule with applications to the Pulmonary Hypertension Association Registry and Colorado COVID-19 hospitalization data
https://bmcmedresmethodol.biomedcentral.com/articles/10.1186/s12874-022-01613-w
-->

<!--
Generalized Additive Models to Capture the Death Rates in Canada COVID-19
https://link.springer.com/chapter/10.1007/978-3-030-85053-1_7 -->

<!--
Contribution of Two-Dose Vaccination Toward the Reduction of COVID-19 Cases, ICU Hospitalizations and Deaths in Chile Assessed Through Explanatory Generalized Additive Models for Location, Scale, and Shape
https://www.frontiersin.org/articles/10.3389/fpubh.2022.815036/full -->

<!-- 
Combining graph neural networks and spatio-temporal disease models to improve the prediction of weekly COVID-19 cases in Germany
https://www.nature.com/articles/s41598-022-07757-5 -->

<!-- 
Association between mobility patterns and COVID-19 transmission in the USA: a mathematical modelling study
https://www.sciencedirect.com/science/article/pii/S1473309920305533 -->

<!-- 
Examining spatiotemporal evolution of racial/ethnic disparities in human mobility and COVID-19 health outcomes: Evidence from the contiguous United States
https://www.sciencedirect.com/science/article/pii/S2210670721007721 -->

<!--
Social, economic, and environmental factors influencing the basic reproduction number of COVID-19 across countries
https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0252373 -->

<!-- 
https://asbates.rbind.io/2019/05/03/gams-for-time-series/
-->

<!--
https://environmentalcomputing.net/statistics/gams/
-->

<!--
https://petolau.github.io/Analyzing-double-seasonal-time-series-with-GAM-in-R/
-->